Formulating RDF Constraints and \\ Validating RDF Data Generically

### 2015-04-21 paper/paper 2/paper 2.tex
paper 2 transformations

\textbf{\emph{Complex constraints}} encompass \emph{simple constraints} and/or \emph{complex constraints}.
\emph{Exclusive or} is a logical operation that outputs true whenever both inputs differ (one is true, the other is false).
The \emph{Context-Specific Exclusive OR of Property Groups}\footnote{Corresponds to R-13-DISJOINT-GROUP-OF-PROPERTIES-CLASS-SPECIFIC} 
is a \emph{complex constraint} restricting individuals of a given context (here a class) to contain only one of multiple property groups.
Publications, e.g., are either identified by an ISBN and a title (for books) or by an ISSN and a title (for periodical publications), but it should not be possible to assign both identifiers to a given publication, which can be represented in ShEx:

@@

As, e.g., the \emph{The-Great-Gatsby} is a \emph{publication} (\ms{rdf:type(The-Great-Gatsby, Publication)}) having an \emph{ISBN} number (\ms{isbn(The-Great-Gatsby, '978-0241965672')}) and a \emph{title} (\ms{title(The-Great-Gatsby, 'The Great Gatsby')}) and not an additional \emph{ISSN} identifier, the data is valid.
%\begin{DL}
%rdf:type(BElannaTorres,Half-Klingon) \\
%klingonMother(BElannaTorres,Miral) \\
%klingonMother(BElannaTorres,JohnTorres)
%\end{DL}
The \emph{complex constraint} above is mapped to DL as follows:

@@

The \emph{complex constraint} (expressed in DL) is mapped to a \emph{generic constraint} using the \emph{RDF-CV} (see table \ref{tab:complex-constraints}).

@@

Even though, tools may generate \emph{generic constraints} automatically, this \emph{complex constraint} is composed of many other \emph{complex} (e.g. minimum and maximum qualified cardinality restrictions) and \emph{simple constraints} (e.g. constraints on sets).
As \emph{exact (un)qualified cardinality restrictions (=n)} and \emph{exclusive or (XOR)} are frequently used \emph{complex constraints},
we propose using them as \emph{simple constraints} - in terms of syntactic sugar.
As a consequence, the \emph{Context-Specific Exclusive OR of Property Groups} constraint above is represented as a \emph{generic constraint} more intuitively and concisely (see table \ref{tab:simplified-complex-constraints}).

@@

\subsection{\emph{Complex Constraints} as \emph{Simple Constraints} (Syntactic Sugar)}

Almost 15 percent of all RDF constraint types are \emph{complex constraints} which can be simplified and therefore formulated as \emph{simple constraints} when using them in terms of syntactic sugar (see section \ref{sec:evaluation}).

There are three forms of \emph{OWL RBox axioms}: \emph{role inclusions}, \emph{equivalence} and \emph{disjointness}. 
OWL provides a variety of others, namely \emph{role transitivity}, \emph{symmetry}, \emph{asymmetry}, \emph{reflexivity} and \emph{irreflexivity}. 
These are sometimes considered as basic axiom types in DL as well, using some suggestive notation such as
\ms{Trans(ancestorOf)} to express that the role \emph{ancestorOf} is transitive. \tb{transitive property example from the library domain}
However, such axioms are just syntactic sugar; 
all role characteristics can be expressed using the basic features of DL.
The \emph{Irreflexive Object Properties} constraint type\footnote{Corresponds to \emph{R-60-IRREFLEXIVE-OBJECT-PROPERTIES}}, e.g., 
restricts that an object property is irreflexive if it is never locally reflexive, i.e. no individual is connected by the object property to itself \cite{Kroetzsch2012}.
The irreflexivity constraint \ms{$\top$ $\sqsubseteq$ $\neg$ $\exists authorOf.Self$}
states that individuals cannot be authors of themselves.
When mapped to the \emph{RDF-CV} (see table \ref{tab:irreflexive-object-properties-as-complex-constraints}), the \emph{complex constraint} aggregates three \emph{simple constraints} (one \emph{property} and two \emph{class constraints}).

@@

When using the OWL RBox axiom \emph{role irreflexivity} as syntactic sugar, 
the \emph{complex constraint} transforms into a \emph{simple property constraint} with exactly the same semantics (see table \ref{tab:irreflexive-object-properties-as-simple-constraints}).

@@

The \emph{Primary Key Properties}\footnote{Corresponds to \emph{R-226-PRIMARY-KEY-PROPERTIES}} constraint type is often useful to declare a given (datatype) property as the "primary key" of a class, so that a system can enforce uniqueness. 
Books, e.g., are uniquely identified by their ISBN, i.e. the property \emph{isbn} is inverse functional \ms{$(\ms{funct } isbn\sp{\overline{\ }})$}, 
which can be mapped to the \emph{RDF-CV} (see table \ref{tab:primary-key-properties-as-complex-constraints}).

@@

\emph{Keys}, however, are even more general, i.e. a generalization of inverse functional properties \cite{Schneider2009}.
A key can be a datatype property, an object property, or a chain of properties.
For these generalization purposes, as there are different sorts of keys, and as keys can lead to undecidability, 
DL is extended with \emph{key boxes} and a special construct \emph{keyfor} \cite{Lutz2005}.
When using the \emph{keyfor} DL construct (\ms{isbn keyfor Book}), the \emph{complex constraint} can be expressed by only one \emph{simple property constraint} (see table \ref{tab:primary-key-properties-as-simple-constraints} for the mapping to the \emph{RDF-CV}).

@@

RDF data providers have to ensure that their data conforms to constraints which are part of explicitly or implicitly concluded contracts between data providers and consumers.
As there is no standard way, these constraints may be represented by a variety of possible constraint languages (there are even more than the five promising ones\footnote{i.a. Bibframe, DQTP, Pellet ICV, RDFUnit, SPARQL, Stardog ICV}).
There is an obvious need to transform any \emph{specific constraint} (\emph{$sc_{\alpha}$}) (expressed by any constraint language \emph{$\alpha$}) into any other \emph{specific constraint} (\emph{$sc_{\beta}$}) (expressed by any constraint language \emph{$\beta$}) - when they have the same semantics.
We use the \emph{RDF-CV} as an intermediate representation vocabulary for these transformations.
If we define mappings ($m(sc_{\alpha}), m(sc_{\beta})$) between equivalent \emph{specific constraints} and the corresponding \emph{generic constraint} (\emph{gc}),
we can convert them automatically:

@@

We do not need to define mappings for each pair of semantically equivalent \emph{specific constraints}.
Let's assume we are able to express a constraint type by 5/6/7/8/9/10 constraint languages.
Without the intermediate \emph{generic constraint}, we would have to define for each constraint type \emph{1} mapping for each pair of \emph{specific constraints} (expressed by different constraint languages)
- that are \emph{\( {n \choose 2} \)} mappings (10/15/21/28/36/45 mappings).
With an intermediate \emph{generic constraint}, however, we only need to define for each constraint type \emph{n} mappings (one mapping for each \emph{specific constraint}) from \emph{n} \emph{specific constraints} to the corresponding \emph{generic constraint} (5/6/7/8/9/10 mappings).

@@

The {\em existential quantification}\footnote{Corresponds to {\em R-86-EXISTENTIAL-QUANTIFICATION-ON-PROPERTIES}}
\ms{Publication $\equiv$ $\exists$ publisher.Publisher} restricts publications to have at least one publisher and can be expressed by at least 10 constraint languages\footnote{at least by OWL2, SPARQL, ShEx, Bibframe, DSP, ReSh, SPIN, DQTP, Pellet IVC, Stardog ICV}:

@@

First, the DL statement \ms{Publication $\equiv$ $\exists$ publisher.Publisher} is mapped to a \emph{simple generic constraint} according to the \emph{RDF-CV} (see table \ref{tab:existential-quantification-as-simple-constraints}).

@@

The next step is to define one mapping for each \emph{specific constraint} (\emph{$sc_{OWL 2}$}, \emph{$sc_{ReSh}$}, \emph{$sc_{ShEx}$}, ...) to the \emph{generic constraint}.

>>>>

\textbf{\emph{Complex constraints}} are composed out of \emph{simple constraints} and/or \emph{complex constraints}.
\emph{Exclusive or} is a logical operation that outputs true whenever both inputs differ (one is true, the other is false).
The \emph{complex constraint} \emph{context-specific exclusive or of property groups}\footnote{Corresponds to R-13-DISJOINT-GROUP-OF-PROPERTIES-CLASS-SPECIFIC} 
restricts individuals of given classes to have properties of only one of multiple mutually exclusive property groups.
\emph{Publications}, e.g., are either identified by an \emph{ISBN} and a \emph{title} (for books) or by an \emph{ISSN} and a \emph{title} (for periodical \emph{publications}), but it should not be possible to assign both identifiers to a given \emph{publication}, which can be expressed in \emph{ShEx} as follows:

@@

As the \emph{The-Great-Gatsby} is a \emph{publication} (\ms{rdf:type(The-Great-Gatsby, Publication)}) having an \emph{ISBN} number (\ms{isbn(The-Great-Gatsby, '978-0241965672')}) and a \emph{title} (\ms{title(The-Great-Gatsby, 'The Great Gatsby')}) and not an additional \emph{ISSN} identifier, the \emph{The-Great-Gatsby} is a valid \emph{publication}.
%\begin{DL}
%rdf:type(BElannaTorres,Half-Klingon) \\
%klingonMother(BElannaTorres,Miral) \\
%klingonMother(BElannaTorres,JohnTorres)
%\end{DL}
The \emph{complex constraint} is mapped to the \emph{RDF-CV} (see table \ref{tab:complex-constraints}) and expressed in \emph{DL} as follows:

@@

@@

%Even though, tools may generate \emph{generic constraints} automatically, this 
\emph{Complex constraints} are composed of many other \emph{complex} (e.g. \emph{minimum and maximum qualified cardinality restrictions}) and \emph{simple constraints} (e.g. constraints on sets).
As \emph{exact (un)qualified cardinality restrictions (=n)} and \emph{exclusive or (XOR)} are frequently used \emph{complex constraints},
we propose using them as \emph{simple constraints} - in terms of syntactic sugar.
As a consequence, the \emph{context-specific exclusive or of property groups} constraint above is represented as a \emph{generic constraint} more intuitively and concisely (see table \ref{tab:simplified-complex-constraints}).

@@

\subsection{\emph{Complex Constraints} as \emph{Simple Constraints}}

Almost 15 percent of all constraint types are \emph{complex constraints} which can be simplified and therefore formulated as \emph{simple constraints} when using them in terms of syntactic sugar (see section \ref{sec:evaluation}).

There are three forms of \emph{OWL RBox axioms}: \emph{role inclusions}, \emph{equivalence} and \emph{disjointness}. 
\emph{OWL} provides a variety of other axiom types: \emph{role transitivity}, \emph{symmetry}, \emph{asymmetry}, \emph{reflexivity} and \emph{irreflexivity}. 
These axiom types are sometimes considered as basic axiom types in \emph{DL} - using some suggestive notation such as
\ms{Trans(ancestorOf)} to express that the role \emph{ancestorOf} is transitive.
Such axioms, however, are just syntactic sugar - 
all role characteristics can be expressed using the basic features of \emph{DL}.
The \emph{irreflexive object properties} constraint type\footnote{Corresponds to \emph{R-60-IRREFLEXIVE-OBJECT-PROPERTIES}} 
restricts that no individual is connected by a given object property to itself \cite{Kroetzsch2012}.
With the following \emph{irreflexive object property} constraint, e.g., one can state that individuals cannot be authors of themselves:
\begin{DL}
\ms{$\top$ $\sqsubseteq$ $\neg$ $\exists authorOf.Self$}
\end{DL}
When mapped to the \emph{RDF-CV} (see table \ref{tab:irreflexive-object-properties-as-complex-constraints}), the \emph{complex constraint} aggregates three \emph{simple constraints} (one \emph{property} and two \emph{class constraints}):

@@

When using the \emph{OWL RBox axiom} \emph{role irreflexivity} in terms of syntactic sugar, 
the \emph{complex constraint} can be expressed more concisely in form of a \emph{simple property constraint} with exactly the same semantics (see table \ref{tab:irreflexive-object-properties-as-simple-constraints}):

@@

The \emph{primary key properties}\footnote{Corresponds to \emph{R-226-PRIMARY-KEY-PROPERTIES}} constraint type is often useful to declare a given (datatype) property as the "primary key" of a class, so that a system can enforce uniqueness. 
\emph{Books}, e.g., are uniquely identified by their \emph{ISBN} number, i.e., the property \emph{isbn} is inverse functional \ms{$(\ms{funct } isbn\sp{\overline{\ }})$}, 
which can be represented by the \emph{RDF-CV} in form of a \emph{complex constraint} (see table \ref{tab:primary-key-properties-as-complex-constraints}):

@@

\emph{Keys}, however, are even more general, i.e., a generalization of \emph{inverse functional properties} \cite{Schneider2009}.
A key can be a datatype, an object property, or a chain of properties.
For these generalization purposes, as there are different sorts of keys, and as keys can lead to undecidability, 
\emph{DL} is extended with \emph{key boxes} and a special construct \emph{keyfor} \cite{Lutz2005}.
When using the \emph{keyfor} \emph{DL} construct (\ms{isbn keyfor Book}), the \emph{complex constraint} can be expressed by only one \emph{simple property constraint} (see table \ref{tab:primary-key-properties-as-simple-constraints} for the mapping to the \emph{RDF-CV}).

@@

RDF data providers have to ensure that their data conforms to constraints which are explicitly defined within contracts between data providers and consumers.
As there is no standard way to express these constraints, they may be represented by a variety of constraint languages - different in syntax and semantics (there are even more than the five promising ones\footnote{i.a. Bibframe, DQTP, Pellet ICV, RDFUnit, SPARQL, and Stardog ICV}).
To overcome the necessity to understand diverse constraint languages, 
we propose to provide a mechanism which transforms any \emph{specific constraint} (\emph{$sc_{\alpha}$}) (expressed by any constraint language \emph{$\alpha$}) into any other \emph{specific constraint} (\emph{$sc_{\beta}$}) (expressed by any constraint language \emph{$\beta$}) - when they have the same semantics.
We use the \emph{RDF-CV} to intermediately represent constraints for these transformations.
By defining mappings ($m(sc_{\alpha}), m(sc_{\beta})$) between equivalent \emph{specific constraints} and the corresponding \emph{generic constraint} (\emph{gc}) we are able to convert them automatically:

@@

We do not need to define mappings for each pair of semantically equivalent \emph{specific constraints}.
Let's assume we are able to express a particular constraint type by 5/6/7/8/9/10 constraint languages.
Without the intermediate \emph{generic constraint}, we would have to define for each constraint type \emph{1} mapping for each pair of \emph{specific constraints} (expressed by different constraint languages)
- that are \emph{\( {n \choose 2} \)} mappings (10/15/21/28/36/45 mappings).
With an intermediate \emph{generic constraint}, however, we only need to define for each constraint type \emph{n} mappings (one mapping for each \emph{specific constraint}) from \emph{n} \emph{specific constraints} to the corresponding \emph{generic constraint} (5/6/7/8/9/10 mappings).

@@

\begin{DL}
Publication $\equiv$ $\exists$ publisher.Publisher
\end{DL}

The {\em existential quantification}\footnote{Corresponds to {\em R-86-EXISTENTIAL-QUANTIFICATION-ON-PROPERTIES}}
%\ms{Publication $\equiv$ $\exists$ publisher.Publisher} 
above restricts \emph{publications} to have at least one \emph{publisher} and can be expressed by at least 10 different constraint languages\footnote{at least by OWL 2, SPARQL, ShEx, Bibframe, DSP, ReSh, SPIN, DQTP, Pellet IVC, and Stardog ICV}:

@@

First, the \emph{DL} statement \ms{Publication $\equiv$ $\exists$ publisher.Publisher} is mapped to a \emph{simple generic constraint} according to the \emph{RDF-CV} (see table \ref{tab:existential-quantification-as-simple-constraints}) and 
the next step is to define one mapping for each \emph{specific constraint} (\emph{$sc_{OWL 2}$}, \emph{$sc_{ReSh}$}, \emph{$sc_{ShEx}$}, ...) to the corresponding \emph{generic constraint}.

@@
